### 2018-05-11 Presentation/Presentation.tex
refining equations

\documentclass{beamer}
\usetheme[
  block=fill,
  background=dark,
  titleformat=smallcaps,
  progressbar=frametitle,
  numbering=none,
]{metropolis}

%----------------------------------------------------------------------------
% LoCo
%----------------------------------------------------------------------------
\setbeamertemplate{frametitle}[default][center]
\newcommand{\xx}{\pmb{\raisebox{-1.5pt}{$\mathcal{X}$}}}
\newcommand{\hh}{\pmb{h}}
\newcommand{\ii}{\pmb{i}}
\newcommand{\oo}{\pmb{o}}
\newcommand{\ff}{\pmb{f}}
\newcommand{\Ss}{\pmb{s}}
\newcommand{\bb}{\pmb{b}}

%Color
\definecolor{mDarkBrown}{HTML}{604c38}
\definecolor{mDarkTeal}{HTML}{23373b}
\definecolor{mLightBrown}{HTML}{EB811B}
\definecolor{mLightGreen}{HTML}{14B03D}
% Math
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{bm}
\newenvironment{rcases}
  {\left.\begin{aligned}}
  {\end{aligned}\right\rbrace}

% Code listing
\usepackage{minted}
%\usemintedstyle{friendly}
\usemintedstyle{tango}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

% Graphs
\usepackage{tikz}
\usetikzlibrary{calc, trees, fit, positioning}
\usepackage{pgfplots}
% Graphics
\usepackage{graphics}
\graphicspath{{figures/}} % Location of the graphics files

\newcommand\todo[1]{\textcolor{red}{#1}}
\newcommand{\w}[1]{\textit{"#1"}}
\newcommand{\sm}[1]{\text{\small{#1}}}

% Box macro
\newcommand{\ex}[2]{
  \vfill
  \begin{alertblock}{#1}
    #2
  \end{alertblock}
}
\newcommand\tsc[1]{\alert{\textsc{#1}}}

%----------------------------------------------------------------------------

% Beamer
\title{Differentiable Neural Computers}
\subtitle{Hybrid Computing using a neural network with dynamic external memory (Graves et al. 2016)}
\author{Konstantinos Kogkalidis}
\date{May 28, 2018}
\institute{Logic and Computation}

\begin{document}
	\maketitle

\begin{frame}{Overview: Probabilistic Programming}
	\alert{Cross-domain}
	\begin{itemize}
	\item Data Flow Programming
	\item Bayesian Reasoning
	\item Machine Learning
	\item Functional Programming
	\end{itemize}
	\pause
	Intuition:
	\textit{"Rather than explicitly write a program, write some \alert{constraints} on the behavior of the desired program and use computational tools to search the program space for \alert{models} satisfying these constraints."}
	\pause
	\[\begin{array}{rcl}
	\textsc{\alert{Program}} &\ & \textsc{\alert{Model}}\\
	\hline
	\text{Discrete} &\ & \text{Continuous}\\
	\text{Deterministic} &\ & \text{Stochastic}\\
	\text{Static} &\ & \text{Adaptive}\\
	\end{array}\]
\end{frame}

\begin{frame}{Overview: DNC}
	\alert{Differentiable Neural Computer}\\
	A recurrent neural network coupled with an external memory.

	\pause
	\begin{itemize}
	\item Inspired by biological memory
	\item Reimaging of classical concepts of computation
	\item Extension of NTMs
	\begin{itemize}
		\item End-to-end differentiable
		\item Auto-associative memory
		\item Turing complete
		\item[+] Stronger memory management
	\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Introduction: Classic Computation}
	\alert{Von Neumann architecture}
	\begin{figure}
	\begin{tikzpicture}
		[auto,
		block/.style ={rectangle, draw=mDarkBrown, thick, fill=gray!20, text width=5em, align=center, rounded corners, minimum height=2em},
		block2/.style ={rectangle, draw=mDarkBrown, thick, fill=mDarkTeal, text width=5em, align=center, rounded corners, minimum height=2em},
		block3/.style ={rectangle, draw=black, thick, fill=gray!20, text width=5em, align=center, rounded corners, minimum height=2em}]
		\draw (-1,0.63) node[block3] (i) {\textcolor{mDarkTeal}{\sc{input}}};
		\draw (6,0.63) node[block3] (o) {\textcolor{mDarkTeal}{\sc{output}}};

		\draw (2.5,-1) node[block2] (m) {\textcolor{gray!20}{\sc{memory}}};
		\draw (2.5,1) node[block] (a) {\textcolor{mDarkTeal}{\sc{alu}}};
		\draw (2.5,2) node[block] (c) {\textcolor{mDarkTeal}{\sc{control unit}}};

		\node[draw=mDarkBrown, thick, fill=mDarkBrown, fit=(a) (c)](Fit1) {};
		\draw (2.5,1) node[block] (a) {\textcolor{mDarkTeal}{\sc{alu}}};
		\draw (2.5,2) node[block] (c) {\textcolor{mDarkTeal}{\sc{control unit}}};

		\node[draw=black, thick, fill=gray!20, fit=(Fit1) (m)](Fit2) {};
		\node[draw=mDarkBrown, thick, fill=mDarkBrown, fit=(a) (c)](Fit1) {};
		\draw (2.5,1) node[block] (a) {\textcolor{mDarkTeal}{\sc{alu}}};
		\draw (2.5,2) node[block] (c) {\textcolor{mDarkTeal}{\sc{control unit}}};
		\draw (2.5,-1) node[block2] (m) {\textcolor{gray!20}{\sc{memory}}};

		\draw (i) edge [->, black, thick] (Fit2);
		\draw (Fit2) edge [->, black, thick] (o);
		\draw ($(m.north) + (0.25,0)$) edge [<-, mDarkBrown, thick] ($(Fit1.south) + (0.25, 0)$ );
		\draw ($(m.north) + (-0.25,0)$) edge [->, mDarkTeal, thick] ($(Fit1.south) + (-0.25, 0)$ );

	\end{tikzpicture}
	\end{figure}
\end{frame}

\begin{frame}{Introduction: Neural Networks}
	\begin{minipage}[t]{.44\textwidth}	
	\alert{Simple Neural Net}
	\begin{figure}[h!]
	\vspace{6.5pt}
	\begin{tikzpicture}[scale=.5]
		\foreach \y /\alph/\name in {
			0/x1/$x_1$,
		  	-1/x2/$x_2$,
		  	-2/x3/$x_3$,
		  	-4/xM/$x_M$
		 }{
		 	\node[circle, draw=gray!90, inner sep=0pt,minimum size=3mm,fill=white] (\alph) at (0,\y) {\tiny \textcolor{mDarkTeal}{\name}};
		 }
		 \node (x..) at (0,-3) {\tiny {\vdots}};
		 \foreach \y /\alph/\name in {
			1/h1/$h_1$,
		  	0/h2/$h_2$,
		  	-1/h3/$h_3$,
		  	-5/hN/$h_N$
		 }{
		 	\node[circle, draw=gray!90, inner sep=0pt,minimum size=3mm,fill=white] (\alph) at (4,\y) {\tiny \textcolor{mDarkTeal}{\name}};
		 }
		 \node (h..) at (4,-3) {\tiny {\vdots}};
		 \foreach \y /\alph/\name in {
			0/y1/$y_1$,
		  	-1/y2/$y_2$,
		  	-2/y3/$y_3$,
		  	-4/yP/$y_P$
		 }{
		 	\node[circle, draw=gray!90,inner sep=0pt,minimum size=3mm,fill=white] (\alph) at (8,\y) {\tiny \textcolor{mDarkTeal}{\name}};
		 }
		 \node (y..) at (8,-3) {\tiny {\vdots}};
		 \foreach \x in {x1, x2, x3, xM}
		 {\foreach \h in {h1, h2, h3, hN}
		 {\draw (\x) edge[gray!90] (\h);}}
		 \foreach \x in {h1, h2, h3, hN}
		 {\foreach \y in {y1, y2, y3, yP}
		 {\draw (\x) edge[gray!90] (\y);}}
	\end{tikzpicture}
	\end{figure}
	\vspace{10pt}
	\centering
	%\small \alert{$y=g(h),\ h=f(x)$}\\
	\small $NN: \pmb{x}_{t_i} \mapsto \pmb{y}_{t_i}$\\
	\alert{No memory}
	\end{minipage}
	\pause
	\hspace{1cm}
	\begin{minipage}[t]{.44\textwidth}	
	\alert{Simple Recurrent Net}
	\begin{figure}[h!]
	\begin{tikzpicture}[scale=.5]
		\foreach \y /\alph/\name in {
			0/x1/$x_1$,
		  	-1/x2/$x_2$,
		  	-2/x3/$x_3$,
		  	-4/xM/$x_M$
		 }{
		% 	\node[] (\alph) at (0,\y) {\name};
		 	\node[circle, draw=gray!90, inner sep=0pt,minimum size=3mm,fill=white] (\alph) at (0,\y) {\tiny \textcolor{mDarkTeal}{\name}};
		 }
		 \node (x..) at (0,-3) {\tiny {\vdots}};
		 \foreach \y /\alph/\name in {
			1/h1/$h_1$,
		  	0/h2/$h_2$,
		  	-1/h3/$h_3$,
		  	-5/hN/$h_N$
		 }{
		 	\node[circle, draw=mLightBrown, inner sep=0pt,minimum size=3mm,fill=gray!20] (\alph) at (4,\y) {\tiny \textcolor{mDarkTeal}{\name}};
		 }
		 \node (h..) at (4,-3) {\tiny {\vdots}};
		 \foreach \y /\alph/\name in {
			0/y1/$y_1$,
		  	-1/y2/$y_2$,
		  	-2/y3/$y_3$,
		  	-4/yP/$y_P$
		 }{
		 	\node[circle, draw=gray!90, inner sep=0pt,minimum size=3mm,fill=white] (\alph) at (8,\y) {\tiny \textcolor{mDarkTeal}{\name}};
		 }
		 \node (y..) at (8,-3) {\tiny {\vdots}};
		 \foreach \x in {x1, x2, x3, xM}
		 {\foreach \h in {h1, h2, h3, hN}
		 {\draw (\x) edge[gray!90] (\h);}}
		 \foreach \x in {h1, h2, h3, hN}
		 {\foreach \y in {y1, y2, y3, yP}
		 {\draw (\x) edge[gray!90] (\y);}}
		 \draw (h1.east) [->, densely dotted, thick, mLightBrown] .. controls +(0.8,0.8) and +(-0.8,0.8) .. (h1.west);
		 \draw (h1.east) [->, densely dotted, thick, mLightBrown] .. controls +(0.4,-0.8) and +(-0.4,0.8) .. (h2.west);
		 \draw (h1.east) [->, densely dotted, thick, mLightBrown] .. controls +(1,-1.6) and +(-0.6,0.4) .. (h3.west);
		 \draw (h1.east) [->, densely dotted, thick, mLightBrown] .. controls +(2,-2) and +(-2,2) .. (hN.west);
	\end{tikzpicture}
	\end{figure}
	\vspace{10pt}
	\centering
	%\small{ \alert{$h(t)=f([x(t); h(t-1)])$}}\\
	\small{ $RNN: \pmb{x}_{t_0} \otimes \pmb{x}_{t_1} \otimes \dots \otimes \pmb{x}_{t_i} \mapsto \pmb{y}_{t_i}$}\\
	\alert{Finite memory}
	\end{minipage}
	\pause

	\begin{quote}
	\small \textit{"If training vanilla neural nets is optimization over functions, training recurrent nets is \alert{optimization over programs}."}
	\end{quote}
\end{frame}

\begin{frame}{Approach}
	Train a RNN to act as the \alert{controller} of a memory matrix M of N 
	addresses through R \alert{read heads} and one \alert{write head}.

	\pause
	\begin{enumerate}
	\item \alert{Content Lookup}\\
		\begin{itemize}
		\item \alert{Attention} over memory defined by weightings $w \in \mathcal{S}^N$
		\item Compare controller output with memory objects (\alert{auto-associative memory})
		\item Allow partial matches  (\alert{pattern completion})
		\end{itemize}
	\pause
	\item \alert{Sequential Retrieval}
		\begin{itemize} 
		\item Fill $L \in [0,1]^{N \times N}$ indexing \alert{temporal transitions}
		\item \alert{Shift} operations defined by $Lw,\ L^\top w$
		\end{itemize}
	\pause
	\item \alert{Dynamic Allocation}
		\begin{itemize}
		\item Mark memory locations with $\{0,1\}$ to \alert{signal usage}
		\item Manipulate signals during R/W operations to enable \alert{reallocation}
		\item Generalization to \alert{unbounded memory}
		\end{itemize}
	\end{enumerate}
\end{frame}

\begin{frame}{Controller: Overview}
	A deep \alert{long short-term memory network} receiving input:
	\[
	\xx_t=[ \pmb{x}_t; \pmb{r}_{t-1}^1; \dots ;\pmb{r}_{t-1}^R] \tag{\text{\alert{timestep t}}}
	\]
	and producing output:
	\[
	(\pmb{\upsilon}_T,\pmb{\xi}_T)=\mathcal{N}([\xx_1;\dots;\xx_T];\vartheta) \tag{\text{\alert{entire sequence}}}
	\]
	where $\mathcal{N}$ a set of state equations and $\vartheta$ their trainable parameters.
\end{frame}

\begin{frame}{Controller: State Equations}
	A more detailed look into $\mathcal{N}$.\\

%	\pause
%	\begin{equation*}
%	\begin{rcases}
%  	\text{\alert{Input}} &: [\xx_t;\hh_{t-1}^l;\hh_t^{l-1}]\\
%  	\text{\alert{Output}} &: \hh_t^l\\
%	\end{rcases}
%	\text{(\alert{Layer \alert{l} at time \alert{t}})}
%	\end{equation*}

	\pause		
	\underline{\alert{LSTM layer equations}}

	\alert{Input}: $[\xx_t;\hh_{t-1}^l;\hh_t^{l-1}]$ \\
	\alert{Output}: $\hh_t^l$
	\pause
	\begin{align}
	\ii_t^l &= \zeta_i([\xx_t;\hh_{t-1}^l;\hh_t^{l-1}]) \tag{\alert{input gate}}\\
	\ff_t^l &= \zeta_f([\xx_t;\hh_{t-1}^l;\hh_t^{l-1}])	\tag{\alert{forget gate}}\\
	\Ss_t^l &= \ff_t^l\Ss_{t-1}^l + \ii_t^l\zeta_s([\xx_t;\hh_{t_1}^l;\hh_t^{l-1}]) \tag{\alert{state update}}\\
	\oo_t^l &= \zeta_o([\xx_t;\hh_{t-1}^l;\hh_t^{l-1}])	\tag{\alert{output gate}}\\
	\hh_t^l &= \oo_t^l\zeta_h(\Ss_t^l)		\tag{\alert{hidden layer}}
	\end{align}
%	\nonumber \\ 
%	\pmb{\upsilon}_t &= W_y[\hh_t^1;\dots;\hh_t^L]		\tag{\alert{output vector}}\\
%	\pmb{\xi}_t &= W_{\xi}[\hh_t^1;\dots;\hh_t^L]		\tag{\alert{interface vector}}
%	\end{align}
%	\begin{align}
%	\ii_t^l &= \sigma(W_i^l[\xx_t;\hh_{t-1}^l;\hh_t^{l-1}]+\bb_i^l)   \tag{\alert{input gate}}\\
%	\ff_t^l &= \sigma(W_f^l[\xx_t;\hh_{t-1}^l;\hh_t^{l-1}]+\bb_f^l)	\tag{\alert{forget gate}}\\
%	\Ss_t^l &= \ff_t^l\Ss_{t-1}^l + \ii_t^ltanh(W_s^l[\xx_t;\hh_{t_1}^l;\hh_t^{l-1}] + \bb_s^l)		\tag{\alert{state}}\\
%	\oo_t^l &= \sigma(W_o^l[\xx_t;\hh_{t-1}^l;\hh_t^{l-1}]+\bb_o^l)		\tag{\alert{output gate}}\\
%	\hh_t^l &= \oo_t^ltanh(\Ss_t^l)		\tag{\alert{hidden}}\\
%	\nonumber \\ 
%	\pmb{\upsilon}_t &= W_y[\hh_t^1;\dots;\hh_t^L]		\tag{\alert{output vector}}\\
%	\pmb{\xi}_t &= W_{\xi}[\hh_t^1;\dots;\hh_t^L]		\tag{\alert{interface vector}}
%	\end{align}
\end{frame}

\begin{frame}{Controller: Signal-Flow (1/2)}
	\alert{Single LSTM layer}
	\begin{figure}
	\begin{tikzpicture}[scale=0.5]
		%-------------------------------------------------------------------------------
		% LSTM
		%-------------------------------------------------------------------------------
		\node[circle, inner sep=0pt,minimum size=3mm,fill=white] (conc) 
			at (-0,6) {\textcolor{mDarkTeal}{$;$}};

		% outer box
		\draw[densely dotted, thick, gray!90] (18,6) -| (18,12);
		\draw[densely dotted, thick,->, gray!90] (18,12) -| (conc);
		%\draw[thick] (0,0) |- (18,12);
		%\draw[thick] (18,12) |- (0,0);

		% core nodes
		\node[circle, draw=black, inner sep=0pt,minimum size=3mm,fill=white] (conc) 
			at (-0,6) {\textcolor{black}{$;$}};
		\node[circle, draw=black, thick, inner sep=0pt,minimum size=6mm,fill=white] (s) 
			at (9,6) {\large\textcolor{black}{$\Ss_t^l$}};
		\node[circle, draw=black, inner sep=0pt,minimum size=3mm,fill=white] (fx) 
			at (9,4) {\textcolor{black}{$\times$}};
		\node[circle, draw=black, thick, inner sep=0pt,minimum size=6mm,fill=white] (f) 
			at (9,2) {\large\textcolor{black}{$\ff_t^l$}};
		\node[circle, draw=black, thick, inner sep=0pt,minimum size=6mm,fill=white] (i) 
			at (6,8) {\large\textcolor{black}{$\ii_t^l$}};
		\node[circle, draw=black, inner sep=0pt,minimum size=3mm,fill=white] (ix) 
			at (6,6) {\textcolor{black}{$\times$}};
		\node[rectangle, draw=black, inner sep=0pt,minimum size=4mm,fill=white] (id) 
			at (4,8) {\textcolor{black}{$\zeta_i$}};
		\node[rectangle, draw=black, inner sep=0pt,minimum size=4mm,fill=white] (sd) 
			at  (4,6) {\small\textcolor{black}{$\zeta_s$}};
		\node[circle, draw=black, thick, inner sep=0pt,minimum size=6mm,fill=white] (o) 
			at (16,10) {\large\textcolor{black}{$\oo_t^l$}};
		\node[circle, draw=black, inner sep=0pt,minimum size=3mm,fill=white] (ox) 
			at (16,6) {\textcolor{black}{$\times$}};
		\node[rectangle, draw=black, inner sep=0pt,minimum size=4mm,fill=white] (hd) 
			at (12,6) {\small\textcolor{black}{$\zeta_h$}};
		\node[rectangle, draw=black, inner sep=0pt,minimum size=4mm,fill=white] (odd) 
			at (4,10) {\small\textcolor{black}{$\zeta_o$}};
		\node[rectangle, draw=black, inner sep=0pt,minimum size=4mm,fill=white] (fd) 
			at (4,2) {\small\textcolor{black}{$\zeta_f$}};
		% recurrency
		\draw (f) edge[thick] (fd);
		\draw[thick] (fd) -| (2,6);
		\draw[thick] (2,6) |- (odd);
		\draw (odd) edge[->,thick] (o);
		\draw[thick] (2,6) |- (id);
		\draw (id) edge[thick] (i);
		% gates
		\draw (o) edge[->, thick] (ox);
		\draw (i) edge[->, thick] (ix);
		\draw (s) edge[->,bend right=45, thick, densely dotted, gray!90] (fx);
		\draw (s) edge[<-,bend left=45, thick] (fx);
		\draw (f) edge[->, thick] (fx);
		% forward
		\draw (-2,4)[->,thick] -| (conc);
		\draw(-2,6) edge[thick,->] (conc);
		\draw (0,6) edge[thick] (sd);
		\draw (sd) edge[->, thick] (ix);
		\draw (ix) edge[->, thick] (s);
		\draw (s) edge[thick] (hd);
		\draw (hd) edge[->, thick] (ox);
		\draw (ox) edge[->, thick] (20,6);

		% labels
		\node[label=below:$\textcolor{white}{\hh_t^{l-1}}$] (ht1) at (-1,4) {};
		\node[label=above:$\textcolor{white}{\xx_t}$] (x) at (-1,6) {};
		\node[label=above:$\textcolor{white}{\hh_t^l}$] (ht) at (19.5,6) {};
		\node[label=above:$\textcolor{white}{\hh_{t-1}^l}$] (x) at (9,12) {};
		\end{tikzpicture}
		\end{figure}
\end{frame}

\begin{frame}{Controller: Signal-Flow (2/2)}
	\alert{LSTM Network} (multiple layers)
	\begin{figure}
	\begin{tikzpicture}[scale=1]
	\tikzset{miniLSTM/.style={draw=black, fill=white, minimum width=20pt, minimum height=10pt, thick}}
	\tikzset{split/.style={circle, draw=black,inner sep=0pt, minimum size = 3mm, fill=white}}
		\node (dummy_node) at (0,0) {};
		\foreach \name/\x in {x1/2, x2/4, x3/6, xd/8, xL/10} {\node[inner sep=0] (\name) at (\x,0) {};}
		\foreach \name/\x in {h1/3, h2/5, h3/7, hL/10} {\node[inner sep=0] (\name) at (\x, -2) {};}
		\foreach \name/\x in {s1/5, s2/7, sL/10} {\node[split] (\name) at (\x,-3) {$\textcolor{black}{\tiny;}$};}
		\foreach \name/\x/\tag in {L1/2/$LSTM_1$, L2/4/$LSTM_2$, L3/6/$LSTM_3$, LL/10/$LSTM_L$} 
		{\node[miniLSTM] (\name) at (\x,-2) {\textcolor{black}{\tiny\tag}};}
		\node (Ld) at (8,-2) {$\dots$};
		\node (sd) at (8,-3) {$\dots$};
		\draw (dummy_node) -| (LL);
		\draw (L1) edge[->,thick] (L2);
		\draw (L2) edge[->,thick] (L3);
		\draw (L3) edge[->,thick] (Ld);
		\draw (Ld) edge[->,thick] (LL);
		\draw[->,thick] (x1) edge (L1.north);
		\draw[->,thick] (x2) edge (L2.north);
		\draw[->,thick] (x3) edge (L3.north);
		\draw[->,thick] (xL) edge (LL.north);
		\draw[->,thick] (xd) edge (Ld.north);
		%\draw[->,thick] (1,0) |- (s0);
		\draw[->,thick] (h1) |- (s1);
		\draw[->,thick] (h2) edge (s1);
		\draw[->,thick] (h3) edge (s2);
		\draw[->,thick] (s1) edge (s2);
		\draw[->,thick] (s2) edge (sd);
		\draw[->,thick] (sd) edge (sL);
		\draw[->,thick] (LL) edge (sL);		
		\draw[->,thick] (sL) edge (10, -5);
		% labels
		\node[label=above right :$\scriptsize\textcolor{white}{\xx_t}$] (xt) at (0,0) {};
		\node[label=above :$\scriptsize\textcolor{white}{\hh_t^1}$] (xt) at (3,-2) {};
		\node[label=above :$\scriptsize\textcolor{white}{\hh_t^2}$] (xt) at (5,-2) {};
		\node[label=above :$\scriptsize\textcolor{white}{\hh_t^3}$] (xt) at (7,-2) {};
		\node[label=below :$\scriptsize\textcolor{white}{[\hh_t^1;\hh_t^2]}$] (xt) at (6,-3) {};
		\node[label=left :$\scriptsize\textcolor{white}{[\hh_t^1;\hh_t^2;\dots;\hh_t^L]}$] (xt) at (10,-5) {};
	\end{tikzpicture}
	\end{figure}
\end{frame}

\begin{frame}{Controller: Outputs}
	\topskip0pt
	\[
	(\pmb{\upsilon}_t,\pmb{\xi}_t)=\mathcal{N}([\xx_1;\dots;\xx_T];\vartheta)
	\]
	\vspace{5pt}

	\pause
	\alert{Intermediate output} $\pmb{\upsilon}_t = W_y[\hh_t^1;\dots;\hh_t^L]$
	\[
	\pmb{y}_t = \pmb{\upsilon}_t + W_R[\pmb{r}_t^1;\dots;\pmb{r}_t^R] \tag{\alert{Memory-conditioning}}
	\]

	\pause
	\alert{Interface vector} $\pmb{\xi}_t = W_\xi[\hh_t^1;\dots;\hh_t^L]$

	\begin{minipage}{0.49\textwidth}
	\begin{itemize}
		\item Read keys: $\pmb{k}_t^{r,i}$ % \in \mathbb{R}^W$
		\item Read strengths: $\beta_t^{r,i}$ % \in [1,\infty)$
		\item Write key: $\pmb{k}_t^w$ % \in \mathbb{R}^W$	
		\item Write strength: $\beta_t^{w}$ % \in [1,\infty)$
		\item Erase vector: $\pmb{e}_t$ % \in [0,1]^W$	
	\end{itemize}
	\end{minipage}
	\begin{minipage}{0.49\textwidth}
	\begin{itemize}
		\item Write vector: $\pmb{v}_t$ % \in \mathbb{R}^W$
		\item Free gates: $\pmb{\phi}_t^i$ % \in [0,1]$		
		\item Allocation gate: $g_t^a$ % \in [0,1]$
		\item Write gate: $g_t^w$ % \in [0,1]$
		\item Read modes: $\pmb{\pi}_t^i$ % \in \mathcal{S}_3$
	\end{itemize}
	\end{minipage}
%	\begin{equation*}
%	\begin{split}
%	\pmb\xi_t = &[\overbrace{\pmb{k}_t^{r,1};\dots;\pmb{k}_t^{r,R};}^\text{\tiny\alert{read keys}}\
%				\overbrace{\hat{\beta}_t^{r,1};\dots;\hat{\beta}_t^{r,R};}^\text{\tiny\alert{read strengths}}\
%				\overbrace{\pmb{k}_t^w;\ \hat{\beta}_t^w;}^\text{\tiny\alert{write key/strength}} \
%				\overbrace{\hat{\pmb{e}}_t;}^\text{\tiny\alert{erase vector}} \\
%				&\underbrace{\pmb{v}_t;}_\text{\tiny\alert{write vector}} \
%				\underbrace{\hat{\pmb{f}}_t^1;\dots;\hat{\pmb{f}_t^R};}_\text{\tiny\alert{free gates}} \
%				\underbrace{\hat{g}_t^a;\ \hat{g}_t^w;}_\text{\tiny\alert{allocation/write gates}}\
%				\underbrace{\hat{\pmb{\pi}}_t^1;\dots;\hat{\pmb{\pi}}_t^R}_\text{\tiny\alert{read modes}}]
%	\end{split}
%	\end{equation*}
\end{frame}

\begin{frame}{Memory Adressing: Content-Lookup}
	\alert{R read keys} $\pmb{k}^{r,i} \in \mathbb{R}^W,\ i=1\dots R$\\
	\alert{R read strengths} $\beta^{r,i} \in [1,\infty),\ i=1\dots R$\\
	\alert{Write key} $\pmb{k}^{w} \in \mathbb{R}^W$\\
	\alert{Write strength} $\beta^{w} \in [1,\infty)$\\

	\pause
	\alert{Matching function} $\mathcal{D}$ comparing memory contents

	\pause
	\alert{Weighting function} $\mathcal{C}$ normalizing and sharpening matches
	\[
	\mathcal{C}(M,\pmb{k},\beta)[i]= 
	\frac {exp \{\beta\mathcal{D} (\pmb{k},M[i,:])\} }
	{ \sum_{j}{exp \{\beta\mathcal{D} (\pmb{k},M[j,:]) \}}}
	\]
\end{frame}

\begin{frame}{Memory Adressing: R/W}
	\alert{Attention} dictated by weightings $\pmb{w} \in \mathcal{S}^N$\\
	\alert{Erase vector} $\pmb{e}_t \in [0,1]^W$\\
	\alert{Write vector} $\pmb{v}_t \in \mathbb{R}^W$\\

	\alert{Read operations}
	\begin{align*}
	 \pmb{r}_t^i = M_t^\top\pmb{w}_t^{r,i}\\
	\end{align*}

	\pause
	\alert{Write operations}
	\begin{align*}
	M_t = \underbrace{M_{t-1} \circ (\pmb{1} - \pmb{w}_t^w\pmb{e}_t^\top)}_{\text{erased memory}} + 
	\underbrace{\pmb{w}_t^w\pmb{v}_t^\top}_{\text{new write}}
	\end{align*}
\end{frame}

\begin{frame}{Memory Adressing: Dynamic Allocation}
	\alert{Free gates} $\pmb{\phi}_t^i \in [0,1]^W$\\
	\alert{Allocation gate} $g_t^a \in [0,1]$\\
	\alert{Write gate} $g_t^w \in [0,1]$\\

	\alert{"Free list" scheme}
	\begin{align*}
	\pmb{\psi}_t &= \prod\limits_{i=1}^R (\pmb{1} - \pmb{\phi}_t^iw_{t-1}^{r,i}) \tag{\alert{memory retention}}\\
	u_t &= (\pmb{u}_{t-1} + \pmb{w}_{t-1}^w + \pmb{u}_{t-1} \circ \pmb{w}_{t-1}^w) \circ \pmb{\psi}_t \tag{\alert{usage tracking}}\\
	\end{align*}

	\pause
	\alert{Attention shift} 
	\begin{itemize}
	\item Obtain the \alert{allocation vector} $\pmb{a}_t$ by normalizing $\pmb{u}_t$\\
	\item \alert{Shift} $\pmb{w}_t$ by $g_t^a\pmb{a}_t$ and \alert{scale} by $g_t^w$
	\end{itemize}
\end{frame}

\begin{frame}{Memory Adressing: Temporal Linking}
	\alert{Free gates} $\pmb{\phi}_t^i \in [0,1]^W$\\

	\alert{Memory retention}
	\begin{align*}
	\pmb{\psi}_t = \prod\limits_{i=1}^R (\pmb{1} - \pmb{\phi}_t^iw_{t-1}^{r,i}) \\
	\end{align*}

	\pause
	\alert{Usage tracking}
	\begin{align*}
	u_t = (\pmb{u}_{t-1} + \pmb{w}_{t-1}^w + \pmb{u}_{t-1} \circ \pmb{w}_{t-1}^w) \circ \pmb{\psi}_t\\
	\end{align*}

	\pause
	\alert{Allocation vector} $\pmb{a}_t$ given by normalizing and sorting $\pmb{u}_t$
\end{frame}

\begin{frame}{Further Reading}
	\begin{itemize}
	\item \alert{Neural Turing Machines} (Graves, Wayne, Danihelka)
	\item \alert{Entity Networks} (Henaff, Weston, Szlam, Bordes, LeCun)
	\item \alert{End-to-End Memory Networks} (Sukhbaatar, Szlam, Weston, Fergus)	
	\item \alert{Jointly Learning to Align and Translate} (Bahdanau, Cho, Bengio)
	\item \alert{Principles of Probabilistic Programming Languages} (Goodman)
	\item \alert{Backprop as a Functor} (Fong, Spivak, Tuyéras)
	\item \alert{Formal Methods for Probabilistic Programming} (Selsam, Liang, Dill)
	\end{itemize}
\end{frame}

\begin{frame}{Conclusion}
	\alert{Takeaway}\\
	We can automatically infer simple functions over complex data structures, in the form of \alert{probability distributions} just by using examples.\\

	\pause
	\center	\alert{Thank you!}
\end{frame}

\end{document}